Using SMT engine to generate Symbolic Automata\thanks{This work was partially funded by the Associated Team FM4CPS between INRIA and ECNU, Shanghai}

### 2017-03-21 2017-Xudong-SMT/OpenPNets-SMT.tex
completed running exemples, and presentation of algebra

\paragraph{Context}
\TODO{
  Summary of previous works on pNets and open pNets => 1/2 page.\\
  Longterm goals:\\
  - open pNets to represent operators and program skeletons (ref to
  \cite{PDP2015})\\
  - compute semantics in term of open transitions\\
  - compute equivalence (bisimulation of open automata)\\
  - model-check properties of open systems
  }
\paragraph{Contribution}

@@

\section{Running Examples}
Several publications \cite{PDP2015,FORTE2016} already have introduced
many examples of pNets, encoding
operators of various classical process algebras, or more complex
synchronisation structures in distributed or parallel languages.
In this section, we use two process expressions, respectively from CCS
and from Lotos, to illustrate different features of pNets. They will
serve as running examples in the whole paper.

@@

The root node of the tree PN1 is the top level of the pNet
  structure. It acts as the parallel operator. There are two subnets,
  noted PN2 and PN3, each representing one of the prefix
  sub-expression. Their behaviors are synchronized in their parent PN1
  according to a set of synchronization vectors, synchronizing input
  and output action from these two subnets to give out a silent action
  $\tau$ or just making the action become visible on the top level when
  one of the subnets is working alone. The vector syntax is typically
  ``<act, \_ > -> act'', in which ``\_'' means that the corresponding
  subnet is not involved in this synchronisation.
  \TODO{The vectors for communication are wrong, we cannot use ?ch(v)
    and !ch(v), see the PDP'15 and Forte'16 papers. }
  Remark that these 4
  synchronisation vectors correspond faithfully to the usual SOS
  rules of CCS operational semantics.

@@

\subsection{Presentation of algebra}
\TODO{Eric: write an introductory paragraph}

\QIN{Before the encoding, the presentation of the pNet should be given first. We can divide the presentation into two parts:}
\begin{itemize}
	\item Sorts: Constants sets of the algebra, types of the data and actions.
	\item Operators: Operators comes from other algebras,  constructors for the parameterized actions.
\end{itemize}

\paragraph{Sorts}
\QIN{Ownership of the variables and constants in a pNet acting as actions or data is probably various. Sorts are represented by their name. The sort of actions is built by the constant actions. For example, let $CCSAct$ be the only set of actions of the CCS running example. Obviously, $CCSAct$ contains constant actions $l$, $r$ in the pLTSs. Sorts here act as sets of actions for pNet, pLTS or holes. Intuitively, Sorts of data and actions are completely independent. Sorts of the data like a type of them could be integer, boolean or a customized type. Sorts for integer and boolean are implicitly declared.}

\paragraph{Operators}
\QIN{Usually we need introduce the expressions from other algebra when encoding the algebra into pNets. So that is needed to know about the input and output of the operators used in the expressions from those algebra. The input and output should belong to the sorts on the pNet. When there are $k$ sorts in this pNet, the structure of a $n$-ary operator is vector as $< Sort_{i_1}, ... , Sort_{i_n} > \rightarrow Sort_j , i_1 , ... , i_n, j \in [0..k] $.}

\paragraph{Extension of other algebras}
\QIN{Variable, constant are treated as a kind of expression together with the expression using operator. Operators can be categorized into unary operator, binary operator and N-ary operator, that make it possible to present operator from other algebras. In Fig .6, we show the Expression class we improved to enable user to extend other algebras in the pNet. There are classes implementing three kinds of expression extend the original  Expression class. From these three classes, more classes are extended using the operators coming from other algebra such as CCS and LOTOS}

@@

@@

\section{Implementation}
\QIN{
VerCors platform uses pNet as the intermediate language for some language or graphical formalism to be translated into both input for a checker and for generating code automatically to ease the specification from non-expert users\cite{FASE}, compiling it from those language or graphical formalism in the pNet API. We compute the open automaton of an open pNet also in that pNet API directly implementing the sketch showed before. However, we made some modification, merging the step 2a-2b and applying Tr2 with the premisses from subnets to generate open transitions, application of Tr1 is still conducted in a simple way. At same time, we apply the fresh function to rename all the variable to make them unique and readable. Besides that, we also have a management of assignments throughout the whole computation. Satisfiability check is done only at top level of the deduction tree construction. An alternative would be to submit the satisfiability check to the SMT solver at each level of the tree construction, potentially reducing the overall number of combinations. But the submission to the SMT engine is costly, and more complexity analysis is required before deciding if this would be worthwhile.
}

@@

\ERIC{
=> Context = use the existing software architecture from VCE (ref to \cite{FASE} paper)\\
=> Current prototype: programatic creation of pNet objects, through the pNet API in VCE. On the middle term, this should be compiled from some language or graphical formalism.\\
=> Open transitions are computed as a direct implementation of the algorithm above. Steps 2a-2b are merged: at each pNet node, after application of rule Tr2, the resulting OT is built from the premisses, without explicitly constructing the deduction tree. \\
=> Variable names are generated for all ``fresh'' and ``clone''
operations, using structured variable names ensuring uniqueness of
fresh names, but also reasonable readability of these names for
debugging purposes.\\
=> Assignments\\
=> Satisfiability check is done only at top level of the deduction tree construction. An alternative would be to submit the satisfiability check to the SMT solver at each level of the tree construction, potentially reducing the overall number of combinations. But submission to the SMT engine is costly, and more complexity analysis is required before deciding if this would be worthwhile.\\
}

>>>>

\paragraph{Context}
\TODO{
  Summary of previous works on pNets and open pNets => 1/2 page.\\
  Longterm goals:\\
  - open pNets to represent operators and program skeletons (ref to
  \cite{HMZ:PDP15})\\
  - compute semantics in term of open transitions\\
  - compute equivalence (bisimulation of open automata)\\
  - model-check properties of open systems
  }
\paragraph{Contribution}

@@

\section{Running Examples}
Several publications \cite{HMZ:PDP15,henrio:Forte2016} already have introduced
many examples of pNets, encoding
operators of various classical process algebras, or more complex
synchronisation structures in distributed or parallel languages.
In this section, we use two process expressions, respectively from CCS
and from Lotos, to illustrate different features of pNets. They will
serve as running examples in the whole paper.

@@

The root node of the tree PN1 is the top level of the pNet
  structure. It acts as the parallel operator. There are two subnets,
  noted PN2 and PN3, each representing one of the prefix
  sub-expression. Their behaviors are synchronized in their parent PN1
  according to a set of synchronization vectors, synchronizing input
  and output action from these two subnets to give out a silent action
  $\tau$ or just making the action become visible on the top level when
  one of the subnets is working alone. The vector syntax is typically
  ``<act, \_ > -> act'', in which ``\_'' means that the corresponding
  subnet is not involved in this synchronisation.
  \TODO{The vectors for communication are wrong, we cannot use ?ch(v)
    and !ch(v), the action algebra has changed since PDP'2015, now you
    must remove the ``?'' and ``!'' from the vectors. }
  Remark that these 4
  synchronisation vectors correspond faithfully to the usual SOS
  rules of CCS operational semantics.

@@

\subsection{Presentation of algebra}
One of the important goal of the Open pNet framework is to provide us
with a model able to express the semantics of many different algebras
or languages. In particular, as we have seen in section
\ref{section:pnets}, the \emph{term algebra} used to express the
action expressions and the data in their parameters is left open in
the pNet formalisation.

So, for a given language, we now need to specify its data and action
domains, giving them an abstract syntax (sorts, constructors, operators,
and predicates). We call this a \emph{Presentation}.
We also define a concrete syntax, that will be used for
pretty-printing, but also for translating the presentation, and the
predicates, into Z3 syntax.

An algebra presentation contains:
\begin{itemize}
	\item Sorts: Constants sets of the algebra, types of the data
          and actions. 
	\item Operators: Operators comes from other algebras \TODO{why
          others ? there can be action constructors also}
          constructors for the parameterized actions.
        \item Constants: The satifiability solver need to know which
          actions in the pLTS behaviours are constants (like the $l$,
          $r$ local actions of the controlers in the CCS example). So
          for a given pNet system, the Action sort will include all
          these constants.
\end{itemize}

\paragraph{Sorts}
Ownership of the variables and constants in a pNet acting as actions
or data is probably various. Sorts are represented by their name. The
sort of actions is built by the constant actions. Sorts here act as
sets of actions for pNet, pLTS or 
holes. There \emph{must} be one sort for actions, and eventually
several others for the types of data parameters. Sorts for integer
and boolean are implicitly declared.

\paragraph{Operators}
We need introduce operators for building expressions, of both action
and data sorts, when encoding the 
algebra into pNets. So that is needed to know about the input and
output of the operators used in the expressions from those
algebra. The input and output should belong to the sorts on the
pNet. When there are $k$ sorts in this pNet, the structure of a
$n$-ary operator is vector as $< Sort_{i_1}, ... , Sort_{i_n} >
\rightarrow Sort_j , i_1 , ... , i_n, j \in [0..k] $.
\TODO{ I do not understand this}

\paragraph{Implementation of the expressions meta-model}
The VerCors platform uses ECORE and EMF to define the structure of
both pNets and actions algebras. There exists a predefined set of
action and data expressions classes, including boolean, integer and
action expressions, with predefined variables, literal (constants),
and operators. Operators can be categorized into unary, binary and
N-ary operators.
The original version of VerCors did not allow to extend these classes,
we have modified its architecture so that a user can extend them,
defining more elements for his specific algebra.

In Fig~\ref{schema:expression-pnets}, we show the Expression class we
improved, including extensions for CCS and LOTOS.

@@

\TODO{give an example of user-defined sorts/operators, and tell that
  you will give it's translation into Z3 in section 5.4}

@@

\section{Implementation}
\QIN{
VerCors platform uses pNet as the intermediate language for some
language or graphical formalism to be translated into both input for a
checker and for generating code automatically to ease the
specification from non-expert users\cite{HKM-FASE16}, compiling it
from those language or graphical formalism in the pNet API. We compute
the open automaton of an open pNet also in that pNet API directly
implementing the sketch showed before. However, we made some
modification, merging the step 2a-2b and applying Tr2 with the
premisses from subnets to generate open transitions, application of
Tr1 is still conducted in a simple way. At same time, we apply the
fresh function to rename all the variable to make them unique and
readable. Besides that, we also have a management of assignments
throughout the whole computation. Satisfiability check is done only at
top level of the deduction tree construction. An alternative would be
to submit the satisfiability check to the SMT solver at each level of
the tree construction, potentially reducing the overall number of
combinations. But the submission to the SMT engine is costly, and more
complexity analysis is required before deciding if this would be
worthwhile. 
}

@@

\ERIC{
=> Context = use the existing software architecture from VCE (ref to \cite{HKM-FASE16} paper)\\
=> Current prototype: programatic creation of pNet objects, through the pNet API in VCE. On the middle term, this should be compiled from some language or graphical formalism.\\
=> Open transitions are computed as a direct implementation of the algorithm above. Steps 2a-2b are merged: at each pNet node, after application of rule Tr2, the resulting OT is built from the premisses, without explicitly constructing the deduction tree. \\
=> Variable names are generated for all ``fresh'' and ``clone''
operations, using structured variable names ensuring uniqueness of
fresh names, but also reasonable readability of these names for
debugging purposes.\\
=> Assignments\\
=> Satisfiability check is done only at top level of the deduction tree construction. An alternative would be to submit the satisfiability check to the SMT solver at each level of the tree construction, potentially reducing the overall number of combinations. But submission to the SMT engine is costly, and more complexity analysis is required before deciding if this would be worthwhile.\\
}
